Probability

### 2016-04-27 prob.tex
def comm proj
prob.tex "Classical Probability Spaces" Add small things from meeting...

\makeatletter
\newtheoremstyle{indented}
  {3pt}% space before
  {3pt}% space after
  {\addtolength{\@totalleftmargin}{3.5em}
   \addtolength{\linewidth}{-3.5em}
   \parshape 1 3.5em \linewidth}% body font
  {}% indent
  {\bfseries}% header font
  {.}% punctuation
  {.5em}% after theorem header
  {}% header specification (empty for default)
\makeatother
\theoremstyle{indented}
%% \theoremstyle{remark}
\newtheorem{example}{Example}
\newtheorem{definition}{Definition}
\newtheorem{cor}{Corollary}
\newcommand{\events}{\ensuremath{\mathcal{E}}}
\newcommand{\qevents}{\ensuremath{\mathcal{E}}}
\newcommand{\pmeas}{\ensuremath{\mu}}
\newcommand{\Hilb}{\mathcal{H}}
\newcommand{\pr}[2]{\langle #1, #2 \rangle}
\newcommand{\ket}[1]{|#1\rangle}
\newcommand{\bra}[1]{\langle#1|}
\newcommand{\ip}[2]{\langle #1 | #2 \rangle}
\newcommand{\proj}[1]{|#1 \rangle\langle #1 |}
\newcommand{\ps}{\texttt{+}}
\newcommand{\ms}{\texttt{-}}
\newcommand{\poss}{{\mbox{\wesa{possible}}}}
\newcommand{\imposs}{{\mbox{\wesa{impossible}}}}

@@

A \emph{probability
  space}~\cite{inun.425605319950101,GrahamKnuthPatashnik1994,rohatgi2011introduction}
specifies the necessary conditions for reasoning coherently about
collections of uncertain events. It consists of a \emph{sample space}
$\Omega$, a space of \emph{events}~$\events$, and a \emph{probability
  measure}~$\pmeas$. In this paper, we will only consider
\emph{finite} sets of events: we therefore define a sample space
$\Omega$ as an arbitrary non-empty finite set and the space of events
$\events$ as $2^\Omega$, the powerset of $\Omega$. Given the set of
events $\events$, a \emph{probability measure} is a function
$\pmeas : \events \rightarrow [0,1]$ such that:
\begin{itemize}
\item $\pmeas(\Omega) = 1$, and 
\item for a collection $E_i$,  of pairwise disjoint events, 
  $\pmeas(\bigcup E_i) = \sum \pmeas(E_i)$.
\end{itemize}

\begin{example}[Two-coin probability space]
  Consider an experiment that tosses two coins. We have four possible
  outcomes that constitute the sample space
  $\Omega = \{ HH, HT, TH, TT \}$. There are 16 total events including
  for example the event $\{ HH, HT \}$ that the first coin is
  ``heads,'' the event $\{ HT, TH \}$ that the two coins land on
  opposite sides, and the event $\{ HT, TH, TT\}$ that at least one
  coin is tails. Here is a possible probability measure for these
  events:
\[\begin{array}{c@{\qquad\qquad}c}
\begin{array}{rcl}
\pmeas(\emptyset) &=& 0 \\
\pmeas(\{ HH \}) &=& 1/3 \\
\pmeas(\{ HT \}) &=& 0 \\
\pmeas(\{ TH \}) &=& 2/3 \\
\pmeas(\{ TT \}) &=& 0 \\
\pmeas(\{  HH, HT \}) &=& 1/3 \\
\pmeas(\{  HH, TH \}) &=& 1 \\
\pmeas(\{  HH , TT \}) &=& 1/3 
\end{array} & \begin{array}{rcl}
\pmeas(\{  HT, TH \}) &=& 2/3 \\
\pmeas(\{  HT , TT \}) &=& 0 \\
\pmeas(\{  TH , TT \}) &=& 2/3 \\
\pmeas(\{  HH, HT, TH \}) &=& 1 \\
\pmeas(\{  HH, HT, TT \}) &=& 1/3 \\
\pmeas(\{  HH, TH, TT \}) &=& 1 \\
\pmeas(\{  HT, TH, TT \}) &=& 2/3 \\
\pmeas(\{  HH, HT, TH, TT \}) &=& 1
\end{array}
\end{array}\]
The assignment satisfies the two constraints for probability measures:
the probability of the entire sample space is 1, and the probability
of every collection of disjoint events (e.g.,
$\{ HT, TH \} = \{ HT \} \cup \{ TH \}$) is the sum of the individual
probabilities. The probability of collections of non-disjoint events
(e.g., $\{ HT, TH, TT \} = \{ HT, TH \} \cup \{ TH , TT \}$) may add
to something different than the probabilities of the individual
events. It is useful to think that this probability measure is
completely induced by the two coins in question and their
characteristics in the sense that each pair of coins induces a
measure, and each measure must correspond to some pair of coins.
\qed\end{example}

In a strict computational or experimental setting, one may question
the reliance of the definition of probability space on the uncountable
and uncomputable real interval $[0,1]$. This interval includes numbers
like $0.h_1h_2h_3\ldots$ where $h_i$ is 1 or 0 depending on whether
Turing machine $M_i$ halts or not. Such numbers cannot be
computed. This interval also includes numbers like $\frac{\pi}{4}$
which can only be computed with increasingly large resources as the
precision increases.
% \yutsung{Check the meaning of the ``computing $\frac{\pi}{4}$'', because
% Bailey Borwein Plouffe formula can computing the $n$th binary digit of $\pi$ 
% using base 16 directly.}
Therefore, in a resource-aware setting, it is more appropriate to
consider probability measures that map events to a finite set of
elements computable with a fixed set of
resources~\cite{Artstein1972,PuriRalescu1983}. The simplest such set,
and the one we will consider exclusively in this paper, is the set
$\mathscr{L}_{2}=\left\{ \imposs, \poss\right\}$ together with the
operation $\vee$ where $x \vee y=\imposs$ if and only if
$x=y=\imposs$. In relation to the first definition, one can interpret
\imposs\ as the closed interval $[0,0]$, \poss\ as the half-open
interval $(0,1]$, and $\vee$ as the addition of intervals. The
definition of a probability measure in this case is modified as being
a function $\pmeas : \events \rightarrow \mathscr{L}_{2}$ such that:
\begin{itemize}
\item $\pmeas(\Omega) = \poss$, and
\item for a collection $E_i$,  of pairwise disjoint events, $\pmeas(\bigcup E_i) = \bigvee \pmeas(E_i)$.
\end{itemize}

\begin{example}[Two-coin probability space with finite set-valued
  probability measure] Under the new set-valued requirement, the
  probability measure in the first example becomes:
\[\begin{array}{c@{\qquad\qquad}c}
\begin{array}{rcl}
\pmeas(\emptyset) &=& \imposs \\
\pmeas(\{ HH \}) &=& \poss \\
\pmeas(\{ HT \}) &=& \imposs \\
\pmeas(\{ TH \}) &=& \poss \\
\pmeas(\{ TT \}) &=& \imposs \\
\pmeas(\{  HH, HT \}) &=& \poss \\
\pmeas(\{  HH, TH \}) &=& \poss \\
\pmeas(\{  HH , TT \}) &=& \poss \\
\end{array} & \begin{array}{rcl}
\pmeas(\{  HT, TH \}) &=& \poss \\
\pmeas(\{  HT , TT \}) &=& \imposs \\
\pmeas(\{  TH , TT \}) &=& \poss \\
\pmeas(\{  HH, HT, TH \}) &=& \poss \\
\pmeas(\{  HH, HT, TT \}) &=& \poss \\
\pmeas(\{  HH, TH, TT \}) &=& \poss \\
\pmeas(\{  HT, TH, TT \}) &=& \poss \\
\pmeas(\{  HH, HT, TH, TT \}) &=& \poss 
\end{array}
\end{array}\]
\qed\end{example}

@@

\begin{definition}[Projection Operators; Orthogonality;
Commutativity~\cite{10.2307/2308516,Redhead1987-REDINA,peres1995quantum,Swart2013}]
Given a Hilbert space $\Hilb$, a projection operator $P$ is a linear transformation from $\Hilb$
to itself such that $P^2 = P = P^{\dagger}$. Projection operators have the
following properties:\footnote {``Projection'' is sometimes called
``orthogonal projection" or ``self-adjoint projection" to
emphasize $P^{\dagger} = P$~\cite{Maassen2010}.}
\begin{itemize}
\item Projection operators $P_1$ and $P_2$ are orthogonal if $P_1P_2 =
  P_2P_1 = \mathbb{0}$;
\item Projection operators $P_1$ and $P_2$ commute if $P_1P_2 =
  P_2P_1$;
\item If the projections $P_1$ and $P_2$ are orthogonal then $P_1 +
  P_2$ is also a projection;
\item If the projections $P_1$ and $P_2$ commute then $P_1P_2$ is also
  a projection; it is not sufficient in this case for $P_1$ and $P_2$
  to be orthogonal. 
\end{itemize}
\end{definition}

>>>>

\makeatletter
\newtheoremstyle{indented}
  {3pt}% space before
  {3pt}% space after
  {\addtolength{\@totalleftmargin}{3.5em}
   \addtolength{\linewidth}{-3.5em}
   \parshape 1 3.5em \linewidth}% body font
  {}% indent
  {\bfseries}% header font
  {.}% punctuation
  {.5em}% after theorem header
  {}% header specification (empty for default)
\makeatother
\theoremstyle{indented}
%% \theoremstyle{remark}
\newtheorem{example}{Example}
\newtheorem{definition}{Definition}
\newtheorem{cor}{Corollary}
\newcommand{\events}{\ensuremath{\mathcal{E}}}
\newcommand{\qevents}{\ensuremath{\mathcal{E}}}
\newcommand{\pmeas}{\ensuremath{\mu}}
\newcommand{\Hilb}{\mathcal{H}}
\newcommand{\pr}[2]{\langle #1, #2 \rangle}
\newcommand{\ket}[1]{|#1\rangle}
\newcommand{\bra}[1]{\langle#1|}
\newcommand{\ip}[2]{\langle #1 | #2 \rangle}
\newcommand{\proj}[1]{|#1 \rangle\langle #1 |}
\newcommand{\ps}{\texttt{+}}
\newcommand{\ms}{\texttt{-}}
\newcommand{\poss}{{\mbox{\wesa{possible}}}}
\newcommand{\imposs}{{\mbox{\wesa{impossible}}}}
\newcommand{\likely}{{\mbox{\wesa{likely}}}}
\newcommand{\unlikely}{{\mbox{\wesa{unlikely}}}}

@@

A \emph{probability
  space}~\cite{inun.425605319950101,GrahamKnuthPatashnik1994,rohatgi2011introduction}
specifies the necessary conditions for reasoning coherently about
collections of uncertain events. It consists of a \emph{sample space}
$\Omega$, a space of \emph{events}~$\events$, and a \emph{probability
  measure}~$\pmeas$. In this paper, we will only consider
\emph{finite} sets of events: we therefore define a sample space
$\Omega$ as an arbitrary non-empty finite set and the space of events
$\events$ as $2^\Omega$, the powerset of $\Omega$. Given the set of
events $\events$, a \emph{probability measure} is a function
$\pmeas : \events \rightarrow [0,1]$ such that:
\begin{itemize}
\item $\pmeas(\Omega) = 1$, and 
\item for a collection $E_i$,  of pairwise disjoint events, 
  $\pmeas(\bigcup_i E_i) = \sum_i \pmeas(E_i)$.
\end{itemize}

\begin{example}[Two-coin probability space]
  Consider an experiment that tosses two coins. We have four possible
  outcomes that constitute the sample space
  $\Omega = \{ HH, HT, TH, TT \}$. There are 16 total events including
  for example the event $\{ HH, HT \}$ that the first coin is
  ``heads,'' the event $\{ HT, TH \}$ that the two coins land on
  opposite sides, and the event $\{ HT, TH, TT\}$ that at least one
  coin is tails. Here is a possible probability measure for these
  events:
\[\begin{array}{c@{\qquad\qquad}c}
\begin{array}{rcl}
\pmeas(\emptyset) &=& 0 \\
\pmeas(\{ HH \}) &=& 1/3 \\
\pmeas(\{ HT \}) &=& 0 \\
\pmeas(\{ TH \}) &=& 2/3 \\
\pmeas(\{ TT \}) &=& 0 \\
\pmeas(\{  HH, HT \}) &=& 1/3 \\
\pmeas(\{  HH, TH \}) &=& 1 \\
\pmeas(\{  HH , TT \}) &=& 1/3 
\end{array} & \begin{array}{rcl}
\pmeas(\{  HT, TH \}) &=& 2/3 \\
\pmeas(\{  HT , TT \}) &=& 0 \\
\pmeas(\{  TH , TT \}) &=& 2/3 \\
\pmeas(\{  HH, HT, TH \}) &=& 1 \\
\pmeas(\{  HH, HT, TT \}) &=& 1/3 \\
\pmeas(\{  HH, TH, TT \}) &=& 1 \\
\pmeas(\{  HT, TH, TT \}) &=& 2/3 \\
\pmeas(\{  HH, HT, TH, TT \}) &=& 1
\end{array}
\end{array}\]
The assignment satisfies the two constraints for probability measures:
the probability of the entire sample space is 1, and the probability
of every collection of disjoint events (e.g.,
$\{ HT, TH \} = \{ HT \} \cup \{ TH \}$) is the sum of the individual
probabilities. The probability of collections of non-disjoint events
(e.g., $\{ HT, TH, TT \} = \{ HT, TH \} \cup \{ TH , TT \}$) may add
to something different than the probabilities of the individual
events. It is useful to think that this probability measure is
completely induced by the two coins in question and their
characteristics in the sense that each pair of coins induces a
measure, and each measure must correspond to some pair of coins.
In this example, the probability measure is induced by a weighted first coin and
a double-headed second coin.
\qed\end{example}

In a strict computational or experimental setting, one may question
the reliance of the definition of probability space on the uncountable
and uncomputable real interval $[0,1]$. This interval includes numbers
like $0.h_1h_2h_3\ldots$ where $h_i$ is 1 or 0 depending on whether
Turing machine $M_i$ halts or not. Such numbers cannot be
computed. This interval also includes numbers like $\frac{\pi}{4}$
which can only be computed with increasingly large resources as the
precision increases.
% \yutsung{Check the meaning of the ``computing $\frac{\pi}{4}$'', because
% Bailey Borwein Plouffe formula can computing the $n$th binary digit of $\pi$ 
% using base 16 directly.}
Therefore, in a resource-aware setting, it is more appropriate to
consider probability measures that map events to a finite set of
elements computable with a fixed set of
resources~\cite{Artstein1972,PuriRalescu1983}. The simplest such set,
and the one we will consider exclusively in this paper, is the set
$\mathscr{L}_{2}=\left\{ \imposs, \poss\right\}$ together with the
operation $\vee$ where $x \vee y=\imposs$ if and only if
$x=y=\imposs$. In relation to the first definition, one can interpret
\imposs\ as the closed interval $[0,0]$, \poss\ as the half-open
interval $(0,1]$, and $\vee$ as the addition of intervals. The
definition of a probability measure in this case is modified as being
a function $\pmeas : \events \rightarrow \mathscr{L}_{2}$ such that:
\begin{itemize}
\item $\pmeas(\Omega) = \poss$, and
\item for a collection $E_i$,  of pairwise disjoint events, $\pmeas(\bigcup_i
E_i) = \bigvee_i \pmeas(E_i)$.
\end{itemize}

\begin{example}[Two-coin probability space with finite set-valued
  probability measure] Under the new set-valued requirement, the
  probability measure in the first example becomes:
\[\begin{array}{c@{\qquad\qquad}c}
\begin{array}{rcl}
\pmeas(\emptyset) &=& \imposs \\
\pmeas(\{ HH \}) &=& \poss \\
\pmeas(\{ HT \}) &=& \imposs \\
\pmeas(\{ TH \}) &=& \poss \\
\pmeas(\{ TT \}) &=& \imposs \\
\pmeas(\{  HH, HT \}) &=& \poss \\
\pmeas(\{  HH, TH \}) &=& \poss \\
\pmeas(\{  HH , TT \}) &=& \poss \\
\end{array} & \begin{array}{rcl}
\pmeas(\{  HT, TH \}) &=& \poss \\
\pmeas(\{  HT , TT \}) &=& \imposs \\
\pmeas(\{  TH , TT \}) &=& \poss \\
\pmeas(\{  HH, HT, TH \}) &=& \poss \\
\pmeas(\{  HH, HT, TT \}) &=& \poss \\
\pmeas(\{  HH, TH, TT \}) &=& \poss \\
\pmeas(\{  HT, TH, TT \}) &=& \poss \\
\pmeas(\{  HH, HT, TH, TT \}) &=& \poss 
\end{array}
\end{array}\]
Compare to the first example, we can still get the information that the second
coin is double-headed, although we loss the information that the first coin is
weighted.
\qed\end{example}

If we consider set with more values, the probability measure may give us more
information about the coins.

@@

\begin{definition}[Projection Operators; Orthogonality;
Commutativity~\cite{10.2307/2308516,Redhead1987-REDINA,peres1995quantum,Swart2013}]
Given a Hilbert space $\Hilb$, a projection operator $P$ is a linear transformation from $\Hilb$
to itself such that $P^2 = P = P^{\dagger}$. Projection operators have the
following properties:\footnote {``Projection'' is sometimes called
``orthogonal projection" or ``self-adjoint projection" to
emphasize $P^{\dagger} = P$~\cite{Maassen2010}.}
\begin{itemize}
\item Projection operators $P_1$ and $P_2$ are orthogonal if $P_1P_2 =
  P_2P_1 = \mathbb{0}$;
\item Projection operators $P_1$ and $P_2$ commute if $P_1P_2 =
  P_2P_1$;
\item If the projections $P_1$ and $P_2$ are orthogonal then $P_1 +
  P_2$ is also a projection;
\item If the projections $P_1$ and $P_2$ commute then $P_1P_2$ is also
  a projection.
\end{itemize}
\end{definition}
